\section*{Chapter 3 Problems}
\addcontentsline{toc}{section}{Chapter 3 Problems}

\begin{enumerate}
\item Which of the following curves are closed? For each of the {\bf closed} curves below, determine whether the curve is positively or negatively oriented.

\begin{enumerate}[a)]
\item The circle of radius 3 centered at $z_0 = -i + 1$.
\item The semicircle of radius 2 centered at 0 starting at $2$ and ending at $-2$.
\item The quarter-circle of radius $\sqrt{2}$, centered at $i$, and going from $1 + 2i$ to $1$.
\item The closed semicircle of radius $R$ from $1$ to $-1$. (Closed here means that once we reach $-1$, we head back to 1 along a straight line.)
\item The triangle with vertices $1 + i$, $2-i$ and $\frac{1}{i}$ starting at $1 + i$ and moving counterclockwise.
\item The square with vertices $0,1,1+i,$ and $i$, travelled in that order.

\item The curve given first by following the circle of radius $2$ centered at $0$ from $2$ to $-2$, counterclockwise. Then follow the line segment from $-2$ to $-1$. Then follow the circle of radius 1 centered at $0$ around one full rotation, clockwise. Then follow the line segment from $-1$ back to $-2$. And then finish following the circle of radius 2 from $-2$ back to $2$, still going counterclockwise.
\end{enumerate}



\item Let $\gamma_1(t)$ be the curve in 1d) and $\gamma_2(t)$ be the curve in 1f). For each $f(z)$ given below, evaluate $\int_{\gamma_1} f(z)dz$ and $\int_{\gamma_2}f(z)dz$.

\begin{enumerate}[a)]
\item $f(z) = z$
\item $f(z) = z^2$
\item $f(z) = e^z$
\item $f(z) = \frac{1}{z + i}$
\item $f(z) = 1$
\end{enumerate}

You should get $0$ for each of these integrals. Do not use any FTC like results, I want you to do these from the definition of the integral.


\item Recall that for a smooth curve $\gamma(t)$ on $[a,b]$, we defined $(-\gamma)(t)$ by:

$$(-\gamma)(t) = \gamma(a + b - t)$$

Prove that $\int_{-\gamma}f(z)dz = -\int_\gamma f(z)dz$.

\item Suppose $\gamma_1$ and $\gamma_2$ are curves such that $\gamma_2$ starts at the end of $\gamma_1$. Prove that:

$$\int_{\gamma_1 + \gamma_2} f(z)dz = \int_{\gamma_1}f(z)dz + \int_{\gamma_2}f(z)dz$$

\item Suppose $\gamma_1(t)$ is a smooth curve, and $\gamma_2(t)$ traverses $\gamma_1(t)$ backwards. Prove that:

$$\int_{\gamma_1 + \gamma_2} f(z)dz = 0$$

Hint: this should be a 2 line argument.


\item Suppose that $\gamma$ is the circle of radius $\frac{1}{2}$ centered at $0$. For which of the following functions $f(z)$ can we use the Cauchy Integral theorem to conclude that $\int_{\gamma}f(z)dz = 0$?

\begin{enumerate}[a)]
\item $f(z) = z$
\item $f(z) = \Log(z)$
\item $f(z) = \Log(z + 4)$
\item $f(z) = (z^2 - z)^{\frac{1}{2}}$, using the principal square root
\item $f(z) = (z^2 - z)^{\frac{1}{2}}$, using the branch of the square root given by $\arg(z) \in (0,2\pi)$
\item $e^{\Log(z)}$
\end{enumerate}


\item Calculate the following integrals using our $\C$FTC:

\begin{enumerate}[a)]
\item $\int_{\gamma} z^2 + \cos zdz$ where $\gamma$ is any curve from $0$ to $2\pi$.
\item $\int_{\gamma} \cos(z)\sin(z)dz$ where $\gamma$ is any curve from $0$ to $2\pi$.
\item $\int_{\gamma} \frac{1}{z}dz$ where $\gamma$ is a curve from $2$ to $-3i$ that doesn't cross $(-\infty,0]$. (Hint: what is the derivative of $\Log(z)$?
\item $\int_{\gamma} \frac{1}{z}dz$ where $\gamma$ is a curve from $a$ to $b$ that does not cross the ray $\{re^{i\theta}|r \ge 0\}$, where neither $a$ nor $b$ are on the ray. 
\item $\int_{\gamma} \frac{1}{z^2}dz$ where $\gamma$ is a curve from $2$ to $-3i$ such that $\gamma(t) \ne 0$ for any $t$.
\item $\int_{\gamma} \Log(z)dz$ where $\gamma$ is a curve from $2$ to $-3i$ that doesn't cross $(-\infty,0]$ (Hint: what is the antidervative of $\ln x$ over $\R$?)
\end{enumerate}

\item Determine which of the following functions have a primitive on the domain $\C\setminus \{0\}$.

\begin{enumerate}[a)]
\item $z^3 - 2z$
\item $\Log z$
\item $z^{\frac{1}{2}}$ for the principal branch.
\item $\Log(e^z)$
\end{enumerate}


\item For each of the below integrals, determine whether Cauchy's Integral Theorem applies. If not, why?

\begin{enumerate}[a)]
\item $\int_{\gamma} \cos(\frac{1}{z})dz$  where $\gamma$ is any curve from $1$ to $i$
\item $\int_{\gamma} \cos(\frac{1}{z})dz$ where $\gamma$ is the circle of radius 1 centered at $1$
\item $\int_{\gamma} \cos(\frac{1}{z})dz$ where $\gamma$ is the circle of radius 1 centered at 2
\item $\int_{\gamma} e^{2z}\Log(z) dz$ where $\gamma$ is the circle of radius 1 centered at $0$
\item $\int_{\gamma} e^{2z}\Log(z) dz$ where $\gamma$ is the circle of radius 1 centered at $2$
\item $\int_{\gamma} e^{2}dz$ where $\gamma$ is the close semicicle of radius 1 centered at $1$ from travelling from $2$ to $0$ clockwise.
\end{enumerate}


\item Determine which of the following domains are simply connected. (Hint: draw it out and look for holes.)

\begin{enumerate}[a)]
\item $\{z \in \C| \RE(z) > 0\}$ 
\item $\{z \in \C| \IM(z) > 0\}$
\item $\{z \in \C| 0 < |z - 2| < 2\}$
\item $\C$
\item $\emptyset$
\item $\C \setminus \Z$
\item open balls
\end{enumerate}


\item Let $D$ be a simply connected domain, and $\gamma_1$ and $\gamma_2$ any two smooth curves in $D$ from $z_0$ to $z_1$. Suppose $f(z)$ is analytic on $D$. Prove that $\int_{\gamma_1} f(z)dz = \int_{\gamma_2}f(z)dz$.

\item Find all isolated singularities of the following functions:

\begin{enumerate}[a)]
\item $z^2 + 1$
\item $\frac{1}{z^2+1}$
\item $e^{\frac{1}{z^2 + 1}}$
\item $\frac{z^3 + 27}{z- \frac{3}{2} - \frac{3\sqrt{3}}{2}i}$
\end{enumerate}

\item For each of the functions above, determine which of their singularities are removable.

\item Which of the following functions have roots inside $|z| = 2$? (Remember that a root of $f$ is a $z\in \C$ with $f(z) = 0$.) If so, what are they?

\begin{enumerate}[a)]
\item $z^3 - \frac{1}{2}$
\item $\sin(4z)$
\item $z^3 - iz^2 - z + i$
\item $e^{4z} - 1$
\item $z^5 - 16$
\end{enumerate}


\item Let $\gamma_1$ be the circle $|z| = 2$ travelled once counterclockwise, and $\gamma_2$ be the circle $|z-1| = 1$ travelled once counterclockwise. For which functions in question 14 is it possible to deform $\gamma_1$ into $\gamma_2$ without crossing any roots of the function? (I.e., for which functions can we use the deformation theorem to conclude that the integrals over these two curves are equal?)


\item Let $\gamma$ be the circle $|z| = 1$ travelled once counterclockwise. Calculate the following integrals using the Cauchy Integral Formula.

\begin{enumerate}[a)]
\item $\int_{\gamma} \frac{1}{z^2}dz$
\item $\int_{\gamma} \frac{1}{z^2 - \frac{1}{4}}dz$
\item $\int_{\gamma} \frac{\cos(z)e^z}{z - \frac{1}{2}}dz$
\item $\int_{\gamma} \frac{\Log(z + 2)}{z}dz$
\item $\int_{\gamma} \frac{1}{z^2 - \frac{5}{2}z + 1}dz$
\item $\int_{\gamma} \frac{1}{(z^2 - \frac{5}{2}z + 1)^2} dz$ for any $n \in \N$
\item $\int_{\gamma} \frac{1}{(z^2 - \frac{5}{2}z + 1)^n} dz$ for any $n \in \N$
\item $\int_{\gamma} \frac{\cos(\sin(\cos(z)))}{z^2}dz$
\end{enumerate}

\item So far we've seen how to handle simple closed curves travelled in positive orientation. What do we do if the curve isn't simple, or the orientation is negative?

\begin{enumerate}[a)]
\item Suppose $\gamma$ can be broken up into $n$ simple closed curves $\gamma_1,...\gamma_n$. How does $\int_{\gamma} f(z)dz$ relate to the integrals $\int_{\gamma_i} f(z)dz$? (No proof required, but it might help to draw a picture.)
\item Suppose $\gamma$ travels a simple closed curve $n$ times. (Meaning that there is some simple closed curve $\Gamma$ such that $\gamma$ traces out $\Gamma$ exactly $n$ times, all in the same direction.)

\vspace{10pt}

Prove that $\int_\gamma f(z)dz = n \int_\Gamma f(z)dz$

\item Suppose $\gamma$ is negatively oriented. Justify why $-\gamma$ is positively oriented. Give a strategy for integrating over negatively oriented curves.
\end{enumerate}


\item Suppose $\Gamma$ is a simple, closed curve. Suppose $\gamma$ travels $\Gamma$ positively $n_1$ times, then negatively $n_2$ times, then positively $n_3$ times, and so on. For $k$ odd, $\gamma$ travels $\Gamma$ positively $n_k$ times, and for each even $k$ it travels $\Gamma$ negatively $n_k$ times. Prove that:

$$\int_{\gamma} f(z)dz = \left(\sum_{k \text{ odd}} n_k - \sum_{k \text{ even}}n_k\right)\int_{\Gamma} f(z)dz$$

\item Let $\gamma$ be the curve defined by travelling $|z| = 1$ twice clockwise, then once counterclockwise, then once clockwise again.

Compute each integral from question 16 over this new curve.

\item Compute each of the following integrals, using any method you like.
\begin{enumerate}[a)]
\item $\int_{|z| = 1} \tan(z)dz$ 
\item $\int_{\gamma} \frac{1}{z^2 - 4z}dz$ over the triangle with vertices $1, 1 - i,$ and $2 + 4i$
\item $\int_{\gamma} \frac{1}{z^2 - 4z}dz$ over the triangle with vertices $-1,1-i,2 + 4i$
\item $\int_{|z-1| = 3} \frac{\cos(z)}{2z^5} dz$
\item $\int_{|z-1| = \frac{1}{2}} \frac{1}{\sin(z)(z-1)^2}dz$
\item $\int_{|z| = 2}\frac{e^z - \sin(z)}{z^2 - 8z + 15}dz$
\item $\int_{|z| = 2} \frac{1}{z^2 -  1}dz$
\item $\int_{|z - 1| = 1} \frac{\sin(z)e^z}{z^4 - 1}dz$
\end{enumerate}

\item  Suppose $f$ is entire with $|f'(z)| < |f(z)|$ for all $z\in \C$. Prove that $f(z) = Ke^{Cz}$ for some $C\in \C$. (Hint: show that $f(z) \ne 0$ for any $z \in \C$. Then, try to use this to come up with a function to use Liouville's theorem on. You may assume that $Ke^{Cz}$ is the only solution to $f'(z) = Cf(z)$.)


\item[\textbf{HARD}] As a preview of things to come, let's see how we can use complex integration to calculate a real integral. We are going to calculate:

$$\int_{-\infty}^\infty \frac{1}{x^4 +1}dx$$

This is an improper integral, which is improper in two places: at $\pm \infty$. So let's recall what this integral is:

$$\int_{-\infty}^\infty \frac{1}{x^4 + 1} dx = \lim_{R \rightarrow \infty} \int_{-R}^0 \frac{1}{x^4 + 1}dx + \lim_{S\rightarrow \infty} \int_0^S \frac{1}{x^4 + 1}dx$$

\begin{enumerate}[a)]

\item Let's start by getting this down to one limit. To begin, show that $\int_{-\infty}^\infty \frac{1}{x^4 + 1}dx$ exists, by comparing it to $\int_{-\infty}^\infty \frac{1}{x^2 + 1}dx$. This shows that we can actually calculate the integral as:

$$\int_{-\infty}^\infty \frac{1}{x^4 + 1}dx = \lim_{R\rightarrow \infty} \int_{-R}^R \frac{1}{x^4 + 1}dx$$

Consider this step optional. It is a good idea to review your first year material on the convergence of improper integrals. We're going to need it in a bit.

\item Let $R > 1$. Now, we can view this as the integral over $\gamma_R(t) = t$ for $t \in [-R,R]$ of:

$$\int_{\gamma_R} \frac{1}{z^4 + 1}dz$$

Now, this doesn't help us much. We have a bunch of techniques for integrating closed curves, but this isn't closed. So let's define a related closed curve.

\vspace{10pt}

Let $C_R$ be the upper semicircle from $R$ to $-R$, so that $\gamma_R + C_R$ is now a closed curve. So our curve is:

\begin{center}\begin{tikzpicture}
\draw[<->] (-2.5,0)--(2.5,0);
\draw[<->](0,-2.5)--(0,2.5);
\draw[-, color = red] (2,0) arc (0:180:2);
\draw[-,color = red] (-2,0) -- (2,0);
\draw[-{Latex[width = 9pt]}, color = red] (2,0) arc (0:160:2);
\draw[-{Latex[width = 9pt]}, color =red] (-2,0)--(1.4,0);
\draw (2,0) node[inner sep = -2pt, label = {[label distance = -1pt]-90:R}]{};
\draw (-2,0) node[inner sep = -2pt, label = {[label distance = -1pt]-90:-R}]{};
\draw (1.4,1.4) node[inner sep = -2pt, label = {[label distance = -1pt]45:$C_R$}]{};
\draw (0.5,0) node[inner sep = -2pt, label = {[label distance = -1pt]-90:$\gamma_R$}]{};
\end{tikzpicture}
\end{center}

Find $\int_{\gamma_R + C_R} \frac{1}{z^4 + 1}dz$.



\item Now, this integral has two components, $\int_{\gamma_R} \frac{1}{z^4 + 1}dz$, which is the integral we care about, and $\int_{C_R} \frac{1}{z^4 + 1}dz$. We'd like to get rid of this second integral. Let's see what happens as we let $R \rightarrow \infty$.

To do this, let's try to estimate this curve. We do this in stages.

\begin{enumerate}[i)]
\item Show that on the curve $|z| = R$, that ${|z^4 + 1|}\ge R^4 - 1$. Use this to show that $\left| \frac{1}{z^4 + 1}\right| \le \frac{1}{R^4 - 1}$.
\item Find the length of $C_R$.
\item Use our estimation of integrals to show that $|\int_{C_R} \frac{1}{z^4 + 1}dz| \le \frac{\pi R}{R^4 - 1}$.
\end{enumerate}

\item Prove that $\lim_{R\rightarrow \infty} \int_{C_R} \frac{1}{z^4 + 1}dz  = 0$.

\item Now, we know $\int_{\gamma_R + C_R} \frac{1}{z^4 + 1}dz$ is constant for $R > 1$, so $\lim_{R\rightarrow \infty} \int_{\gamma_R + C_R} \frac{1}{z^4 + 1}dz$ exists.

\vspace{10pt}

We also know that $\lim_{R\rightarrow 0} \int_{C_R}\frac{1}{z^4 + 1} dz = 0$.

\vspace{10pt}

Use these two facts to find $\lim_{R\rightarrow \infty} \int_{\gamma_R} \frac{1}{z^4 + 1}dz$.

\item Show that $\int_{-\infty}^\infty \frac{1}{x^4 + 1}dx = \frac{\pi}{\sqrt{2}}$.

\end{enumerate}
\end{enumerate}

